---
title: "Reproducible Research: Peer Assessment 1"
author: "Yanyuan Zhu"
date: Aug 13, 2020
output: md_document
#  html_document:
#    keep_md: true
---
## Load packages and set work directory
```{r echo = TRUE}
## load packages
library(data.table)
library(ggplot2)
# Fork github url before setting wd
# set work directory
setwd("~/Soft/Rtest/RepData_PeerAssessment1")
```



## Loading and preprocessing the data

Download the source code file
```{r echo = TRUE}
## Download the source code file
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./data/repdata_data_activity.zip")
```

unzip the file in linux, then
```{r echo = TRUE}
activity <- data.table::fread('./data/activity.csv')
```


## What is mean total number of steps taken per day?
Calculate the total number of steps per day
```{r echo = TRUE}
#activity[, lapply(.SD, sum, na.rm = FALSE)]
total_steps <- activity[, c(lapply(.SD, sum, na.rm = FALSE))
         , .SDcols = c("steps")
         , by = .(date)         
        ]
head(total_steps)
```

Make a histogram of the total number of steps taken each day. 
```{r echo = TRUE}
g <- NULL
g <- ggplot(total_steps, aes(x = steps))
g <- g + geom_histogram(fill = 'green', binwidth = 1000)
g <- g + labs(
    title = "Daily Steps Taken"
    , x = 'Steps'
    , y = "Frequency"
)
g
```

Report mean and median values of the total number of steps taken per day

```{r echo = TRUE}
total_steps[, .(Mean_steps = mean(steps, na.rm = TRUE)
               , Median_steps = median(steps, na.rm = TRUE)
               )
           ]
```

## What is the average daily activity pattern?
Make a dataframe named *interval*, showing average number of steps taken everyday

```{r echo = TRUE}
interval <- activity[
    , c(
        lapply(.SD, mean, na.rm = TRUE)
       )
    , .SDcols = c("steps")
    , by = .(interval)
]
head(interval)
```

Plot the average steps (y-axis) and days (x-axis)
```{r echo = TRUE}
g <- ggplot(interval, aes(x = interval, y = steps))
g <- g + geom_line(color = 'red', size = 1)
g <- g + labs(
    title = "Average Daily Steps"
    , xlab = 'Interval'
    , ylab = "Average Steps Per Day"
)
g
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo = TRUE}
interval[steps == max(steps), .(max_interval = interval)]
```
## Imputing missing values

Check number of NAs in the column steps
```{r echo = TRUE}
#activity[is.na(steps), .N]
nrow(activity[is.na(steps), ])
```

Replace NAs with median values
```{r echo = TRUE}
#activity[is.na(steps), "steps"] <- 
activity[is.na(steps), 'steps'] <- activity[, c(lapply(.SD, median, na.rm=TRUE))
                                             , .SDcols = c("steps")
                                            ]
```

Export a tidy dataset with missing value filled

```{r echo = TRUE}
data.table::fwrite(x = activity, file = './data/tidyData.txt', quote = FALSE)
```

Total number of steps taken per day
```{r echo = TRUE}
# total number of steps taken per day
total_steps <- activity[, c(lapply(.SD, sum)), .SDcols = c("steps"), by = .(date)]
head(total_steps)
```
Mean and median total number of steps taken per day
```{r echo = TRUE}
# mean and median total number of steps taken per day
total_steps[, .(mean_steps = mean(steps), median_steps = median(steps))]
```

Make the plot
```{r echo = TRUE}
g <- NULL
g <- ggplot(total_steps, aes(x = steps))
g <- g + geom_histogram(fill = "darkblue", binwidth = 1000)
g <- g + labs(
    title = "daily Steps Taken"
    , x = "Steps"
    , y = "Frequency"
)
g
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor column with 2 levels: weekday and weekend
```{r echo = TRUE}
# create a new factor column with 2 levels: weekday and weekend
activity <- fread(input = './data/activity.csv')
activity[, date := as.POSIXct(date, format = "%Y-%m-%d")]
activity$"Day_of_Week" = weekdays(activity$date)
```
Create a new column, indicating whether the day is weekday (Monday to Friday) or weekend (Saturday, Sunday)
```{r echo = TRUE}
activity [grepl(pattern = "æ˜ŸæœŸä¸€|æ˜ŸæœŸäºŒ|æ˜ŸæœŸä¸‰|æ˜ŸæœŸå››|æ˜ŸæœŸäº”", x = activity$'Day_of_Week'), "Weekday_or_Weekend"] <- "Weekday"
activity [grepl(pattern = "æ˜ŸæœŸå…­|æ˜ŸæœŸæ—¥", x = activity$'Day_of_Week'), "Weekday_or_Weekend"] <- "Weekend"
```

Check the dataframe
```{r echo = TRUE}
str(activity)
```

Replace missing values in "steps" column by median value
```{r echo = TRUE}
# Replace missing values in "steps" column by median value
activity[is.na(steps), 'steps'] <- median(activity$steps, na.rm = TRUE)
```

Create a new dataframe, replace missing values in "steps" column by mean value, by interval and "Weekday or Weekend"
```{r echo = TRUE}
# Create a new dataframe, replace missing values in "steps" column by mean value, by interval and "Weekday or Weekend"
interval_df <- activity[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `Weekday_or_Weekend`)]
#interval_df$"Weekday_or_Weekend" <- as.factor(interval_df$"Weekday_or_Weekend")
interval_df$Weekday_or_Weekend <- as.factor(interval_df$Weekday_or_Weekend)
```

```{r echo = TRUE}
str(interval_df)
```

Make a panel plot containing a time series plot (i.e. ðšðš¢ðš™ðšŽ = "ðš•") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r echo = TRUE}
g <- NULL
g <- ggplot(interval_df, aes(x = interval, y = steps, color = Weekday_or_Weekend))
g <- g + geom_line()
g <- g + facet_wrap(~Weekday_or_Weekend, nrow = 1, ncol=2)
g
```
